---
title: "Lab3"
subtitle: "Determinants of crime"
author: "David Hou, Scott Hungerford, Irene Seo"
output:
  pdf_document: default
  html_document: default
---

```{r}
setwd("~/Desktop/W203.2/Assignments/Lab_3")
library(dplyr)
library(ggplot2)
library(stargazer)
```

## Introduction
We are hired to examine the data to help the campaign understand the determinants of crime and to generate policy suggestions that are applicable to local government.

### Cleansing Data
First we clean the data, using the dplyr package for its nice verbs. We remove NAs, change prbconv to numeric, and change all integer columns to factors.

```{r}
raw = as_tibble(read.csv('crime_v2.csv'))
t = raw %>% 
    filter(!is.na(county)) %>%
    mutate(prbconv = as.numeric(prbconv) / 100) %>%
    mutate_if(is.integer, as.factor)
levels(t$west) = c('East', 'West')
t$west = relevel(t$west, 'West') # Put West first so it appears on the left on facet plots 
levels(t$central) = c('Outer', 'Central')
levels(t$urban) = c('Non-urban', 'Urban')
```

As a data transformation, we sum up all of the wage types to make a single total wage.

```{r}
t = t %>% mutate(wage = wcon + wtuc + wtrd + wfir + wser + wmfg + wfed + wsta + wloc)
str(t)
```

We have 91 observations from the data set to analyze.

## Univariate Analysis

First we examine variables of interest to a politician with regards to changing policies.
1. Crime rate
2. Tax revenue per capita
3. Wages
4. Police per capita
5. Average sentences in days

```{r}
summary(t$crmrte)
qplot(t$crmrte, geom = 'histogram', col = I('white'), main = 'Crime Rate', xlab = 'Crime Rate')
```

Crime rate shows skewed distribution with mean at 3.3%.

```{r}
summary(t$taxpc)
qplot(t$taxpc, geom = 'histogram', col = I('white'), main = 'Tax Revenue Per Capita', xlab = 'Tax Revenue Per Capita')
```

Tax revenue per capita also shows skewed distribution with mean value at 38 thousand dollars. (?)

```{r}
summary(t$wage)
qplot(t$wage, geom = 'histogram', col = I('white'), main = 'Wages', xlab = 'Wages')
```

Wage also shows slightly skewed distribution similar to that of crime rate, with mean value at 2,955 and with some outliers to the right side of distribution.

```{r}
summary(t$prbarr)
qplot(t$prbarr, geom = 'histogram', col = I('white'), main = 'Probability of arrest', xlab = 'Probability of arrest')
```

Probability of arrest has a left skewed distribution around mean at 29%.


```{r}
summary(t$prbconv)
qplot(t$prbconv, geom = 'histogram', col = I('white'), main = 'Probability of conviction', xlab = 'Probability of conviction')
```

Probability of conviction has a uniform distribution.

```{r}
summary(t$prbpris)
qplot(t$prbpris, geom = 'histogram', col = I('white'), main = 'Probability of prison sentence', xlab = 'Probability of prison sentence')
```

Probability of prison sentence has a right skewed distribution around mean at 41%.

```{r}
summary(t$polpc)
qplot(t$polpc, geom = 'histogram', col = I('white'), main = 'Police per capita', xlab = 'Police per capita')
```

Police per capita has very skewed distribution with mean at .0017. It also has an extreme outlier to the right.


### Crime rates by regions

Next we examine crime rates by regions and by density of population.

```{r}
ggplot(t, aes(crmrte)) + 
    geom_histogram() + 
    facet_grid(west ~ central) + 
    theme(panel.spacing = unit(1, "lines")) +
    labs(title = 'Crime Rate by Region', x = 'Crime Rate')
```

Our data points come more from Eastern N.C. than Western N.C., and more from outer region than central region.
Except for Western central region, they all show left skewed distributions. Eastern N.C. has crime rate spikes around 2.5%, whereas crime rate peaks at 1% in Western outer region.

```{r}
ggplot(t, aes(crmrte)) + 
    geom_histogram() + 
    facet_grid(. ~ urban) + 
    theme(panel.spacing = unit(2, "lines")) +
    labs(title = 'Non-urban vs Urban Crime Rate', x = 'Crime Rate')
```

We have far more data points of crimes in non-urban area than urban area. (?)
Crime rates are densely populated around 3% in non-urban area with a left skewed distribution, whereas urban crime rates are dispersed around higher value of 5% to 10%.

## Bivariate Analysis

```{r}
ggplot(t, aes(crmrte, taxpc)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Tax Revenue per Capita', x = 'Crime Rate')
lm(crmrte ~ taxpc, data=t)
```

```{r}
ggplot(t, aes(crmrte, wage)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Wages', x = 'Crime Rate')
lm(crmrte ~ wage, data=t)
```


```{r}
ggplot(t, aes(crmrte, prbarr)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Arrest Probability', x = 'Crime Rate')
lm(crmrte ~ prbarr, data=t)

ggplot(t, aes(crmrte, prbconv)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Conviction Probability', x = 'Crime Rate')
lm(crmrte ~ prbconv, data=t)

ggplot(t, aes(crmrte, prbpris)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Prison Probability', x = 'Crime Rate')
lm(crmrte ~ prbpris, data=t)

ggplot(t, aes(crmrte, avgsen)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Average Prison Sentence', x = 'Crime Rate')
lm(crmrte ~ avgsen, data=t)

ggplot(t, aes(crmrte, polpc)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(title = 'Crime Rate vs Police Per Capita', x = 'Crime Rate')
lm(crmrte ~ polpc, data=t)
```



